Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-17 paper-2023-acsos-pulverization.tex
fix dangling sentences

Although the idea behind pulverisation is simple,
finding a way to implement it effectively requires tackling several challenges at different levels:
\begin{itemize}
    \item \textbf{communication}: spliting logical devices into small deployment units implies communication among them,
    thus requiring to consider networking at two levels:
    across pulverised components (intra-device) and among logical devices (inter-device);
    \item \textbf{portability}: hosts will have very diverse hardware specifications, operang systems, and software stacks,
    \item \textbf{runtime}: runtime reconf how to handle the communication between pulverised components across different runtimes?
    \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
    the pulverised configuration
    and the mapping of pulverised components to the underlying infrastructure
    can be expressed nimbly and, possibly, declaratively.
\end{itemize}

@@

\section{A \ac{dsl} for runtime reconfiguration of pulverised systems}\label{sec:contribution}

@@

\subsection{Implementation details}

We implemented the proposed \ac{acronym} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/nicolasfara/pulverization-framework}}.

>>>>

Although the idea behind pulverisation is simple,
finding a way to implement it effectively requires tackling several challenges at different levels:
\begin{itemize}
    \item \textbf{communication}: spliting logical devices into small deployment units implies communication among them,
    thus requiring to consider networking at two levels:
    across pulverised components (intra-device) and among logical devices (inter-device);
    \item \textbf{portability}: hosts will have very diverse hardware specifications, operating systems, and software stacks;
    \item \textbf{runtime}: the system should be able to reconfigure its deployment at runtime without disrupting the application logic; and
    \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
    the pulverised configuration
    and the mapping of pulverised components to the underlying infrastructure
    can be expressed nimbly and, possibly, declaratively.
\end{itemize}

@@

\section{A DSL for reconfigurable pulverised systems}\label{sec:contribution}

@@

\subsection{Implementation details and underlying runtime}

We implemented the proposed \ac{dsl} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/nicolasfara/pulverization-framework}}.
